\subsubsection*{Purpose, Description}

GlueX is a compute-intensive experiment and will require a substantial
computing infrastructure at JLab.  This involves:
\begin{itemize}
\item design and implementation of a fast backbone network at the lab
\item design and implementation of system to transfer 100 MB/sec from
  the counting house to the central computing facility
\item purchase and installation of adequate offline storage media
  (approx. 1 PB/year raw data, 2 PB/year generated data)
\item creation and maintenance of large offline analysis farms (hundreds of cpu's)
\item implementation of grid services to allow
transparent access (OC 24 or better) to JLab and university computing
facilities by GlueX collaborators
\end{itemize}



\subsubsection*{Current Status, R\&D issues}

Much of the above will not be needed for a number of years.  Some of
the items involve extension and/or upgrade of existing facilities,
while others are completely new (e.g. grid services).  JLab is a
member of PPDG (a major grid collaboration) and is actively involved
in grid R\&D.  We note that these all efforts are a continuation of
current computer center activities.



\subsubsection*{Collaboration Responsibilities}

Development and maintenance of the systems mentioned above will be the
primary responsibility of the JLab Computing Center, although GlueX
collaborators will need to be involved with the planning effort to
ensure GlueX needs are met.  

We estimate that prototypes of the items above will need to be in
place three to five years before the beginning of data taking, and
that production versions should be working one to three years before
data taking (depending on system).  We reconize that the purchase of
many components (e.g. farm processors, switches, etc.) must be
postponed for as long as possible for economic reasons, and that what
needs to be in place early is the infrastructure within which these
reside.

We note that the CEBAF Center extension currently under construction
has dedicted space to house the offline farms and tape storage
facilities needed to support GlueX.

