\subsubsection*{Purpose, Description}

This effort includes development of all software needed to analyze GlueX
data, and includes:
\begin{itemize}
\item detector calibration 
\item full event reconstruction
\item simulation
\item partial wave analysis
\item creation of data summary files
\item level 3 trigger algorithm
\end{itemize}


\subsubsection*{Current Status, R\&D issues}

GlueX has developed a preliminary simulation suitable for experiment design
and acceptance studies, and has begun work on an offline analysis
framework and on the experiment data model.  GlueX is also a leader in
the world-wide effort to develop new PWA tools and algorithms.  


\subsubsection*{Collaboration Responsibilities}

Most of the GlueX analysis and simulation algorithms and software
remains to be developed.  This is a very large effort that has already
begun, and it will continue for years after experiment startup.  The
work will be done by the GlueX collaboration, and a large fraction of
GlueX members are expected to contribute.

Preliminary calibration software must be ready about a year before
startup, and the final system must be ready by the time the experiment
begins.

A full-featured event simulation, analysis framework, and preliminary
reconstruction and analysis software needs to be available two to four
years before the start of data taking.  Final reconstruction and
analysis software needs to be ready about six months before data
taking for GlueX to succeed in publishing preliminary results within a
year of startup.

The production level 3 trigger algorithm is needed about a year after
startup, as initial data taking will be at a reduced rate, but a
prototype must be available about one year before startup for testing.

